---
title: "Demo"
author: "Adam Hardaker"
date: "2024-06-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(emmeans)
library(psych)
library(car)
library(gridExtra)
library(ggplot2)
```

loading the data
```{r}
setwd("C:/Users/mahar/OneDrive/Academia/Kassel EBGo/2. Summer 24/QMEE/otree_dev/QMEE/analysis")
file_path <- "data/ACTUAL_EXP_CREAMERY_QMEESS24.csv"

# Read the CSV file
df <- read_csv(file_path)
```

Cleaning data and using long format
```{r}
# Extract survey responses
survey_responses <- df %>%
  select(
    participant_id = `participant.id_in_session`,
    participant_label = `participant.label`,
    altruism = `centipede_game.4.player.altruism`,
    trust = `centipede_game.4.player.trust`,
    risk_tolerance = `centipede_game.4.player.risk_tolerance`,
    greed = `centipede_game.4.player.greed`,
    desire_to_win = `centipede_game.4.player.desire_to_win`
  ) %>%
  mutate(across(altruism:desire_to_win, ~as.numeric(as.character(.))))  # Convert to numeric

# Restructure the game data
restructured_data <- df %>%
  pivot_longer(
    cols = starts_with("centipede_game."),
    names_to = c("round", ".value"),
    names_pattern = "centipede_game\\.(\\d+)\\.(.+)",
    values_drop_na = TRUE
  ) %>%
  select(
    participant_id = `participant.id_in_session`,
    participant_label = `participant.label`,
    round,
    group_id = `group.id_in_subsession`,
    role = `player.id_in_group`, # indicates first-mover (1) or second mover (2)
    treatment = `group.treatment`,
    node_reached = `group.node`,
    payoff = `player.payoff`,
    treatment_order = `group.treatment_order`,
    
  ) %>%
  mutate(
    round = as.integer(round),
    group_id = as.integer(group_id),
    role = as.integer(role),
    node_reached = as.integer(node_reached),
    payoff = as.numeric(payoff),
  )

# Join game data with survey responses
final_data <- restructured_data %>%
  left_join(survey_responses, by = c("participant_id", "participant_label"))

# Sort the data by group_id, then round, then role
sorted_data <- final_data %>%
  arrange(group_id, round, role)

rm(restructured_data,final_data,df)

data<-sorted_data
# you can also save this if you wanna
```

## Analysis

Descriptive stats:
1. Sample size and demographics
2. Number of groups
3. Treatment-Specific Statistics
4. Round specific statistics
5. Treatment Order Effects
6. Personality Traits
7. Group dynamics
- frequency of games ended due to player 1 vs player 2 decisions
- average difference in payoffs between player 1 and player 2

Checking for balance: bar chart and chi-squared
```{r}
# Count the number of games played in each treatment
treatment_counts <- data %>%
  group_by(treatment) %>%
  summarise(
    num_games = n_distinct(group_id, round),
    num_participants = n_distinct(participant_id)
  )

# Print the results
print(treatment_counts)

# Visualize the balance
ggplot(treatment_counts, aes(x = treatment, y = num_games, fill = treatment)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = num_games), vjust = -0.5) +
  labs(title = "Number of Games Played per Treatment",
       x = "Treatment",
       y = "Number of Games") +
  theme_minimal()

# Chi-square test for equal proportions
chisq_result <- chisq.test(treatment_counts$num_games)
print(chisq_result)
```


```{r}
# Overall game statistics
overall_stats <- data %>%
  summarise(
    mean_node = mean(node_reached),
    median_node = median(node_reached),
    sd_node = sd(node_reached),
    mean_payoff = mean(payoff)
  )

# Treatment-specific statistics
treatment_stats <- data %>%
  group_by(treatment) %>%
  summarise(
    mean_node = mean(node_reached),
    median_node = median(node_reached),
    sd_node = sd(node_reached),
    mean_payoff = mean(payoff)
  )

# Round-specific statistics
round_stats <- data %>%
  group_by(round) %>%
  summarise(
    mean_node = mean(node_reached)
  )

# Personality traits summary
personality_stats <- data %>%
  select(altruism, trust, risk_tolerance, greed, desire_to_win) %>%
  describe()

# Correlation of traits with game outcomes
trait_correlations <- data %>%
  select(node_reached, payoff, altruism, trust, risk_tolerance, greed, desire_to_win) %>%
  cor()

# Print results
print(overall_stats)
print(treatment_stats)
print(round_stats)
print(personality_stats)
print(trait_correlations)

# You might also want to create visualizations, e.g.:
ggplot(data, aes(x = treatment, y = node_reached)) +
  geom_boxplot(varwidth=T, fill="plum", color="black") + 
    labs(title="Node Reached Grouped by Treatment",
         x="Treatment",
         y="Node Reached",
         fill="Treatment") +
  theme_minimal()
```

Average Nodes Reached by Round and Treatment
```{r}
ggplot(data, aes(x = round, y = node_reached, color = treatment)) +
  geom_line(stat = "summary", fun = "mean", linewidth = 2) +
  labs(title = "Average Nodes Reached by Round and Treatment",
       y="Node Reached",
       x="Round",
       color="Treatment") +
  theme_minimal() +
  theme(legend.position = c(0.8, 0.8),  # Adjust these values as needed
        legend.background = element_rect(fill = "white", color = "grey"))

```

test normality
```{r}
shapiro_test <- shapiro.test(data$node_reached)
shapiro_test
```
Data significantly deviates from normal distribution

Histogram by treatment
```{r}
ggplot(data, aes(x = factor(node_reached), fill = treatment)) +
  geom_bar(position = "dodge", stat = "count") +
  scale_fill_manual(values = c("control" = "#F8766D", "higher_fixed" = "#00BA38", "higher_random" = "#619CFF")) +
  labs(x = "Node Reached", 
       y = "Frequency", 
       fill = "RoundType") +
  theme_minimal() +
  theme(#panel.grid.major = element_line(color = "gray90"),
        #panel.grid.minor = element_blank(),
        #panel.background = element_rect(fill = "white"),
        #axis.line = element_line(color = "black"),
        legend.position = c(0.85, 0.85),  # Position legend inside the plot
        legend.background = element_rect(fill = "white", color = "gray", size = 0.5),
        legend.margin = margin(6, 6, 6, 6),
        legend.key.size = unit(1, "lines")) +  # Make legend keys smaller
  scale_x_discrete(limits = factor(1:6)) +  # Ensure all nodes from 1 to 6 are shown
  scale_y_continuous(expand = c(0, 0), limits = c(0, 45))  # Adjust y-axis to start at 0 and extend to 45
```
Doesn't look normal

```{r}

# Poisson regression for count data (if `node_reached` is count-like)
glm_model <- glm(node_reached ~ treatment + round, family = "poisson", data = data)


# Extract residuals
residuals <- residuals(glm_model)


# Q-Q Plot for Residuals
ggplot(data.frame(residuals = residuals), aes(sample = residuals)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Q-Q Plot of Residuals",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()
```


## statistical tests

Use wide data for Friedman test
```{r}
data_wide <- data %>%
  group_by(participant_id, treatment) %>%
  summarize(avg_node_reached = mean(node_reached, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = avg_node_reached) %>%
  drop_na()

print(head(data_wide))

friedman_result <- friedman.test(as.matrix(data_wide[, c("control", "higher_fixed", "higher_random")]))
print(friedman_result)
```

Check learning effects
```{r}
# Prepare data for learning effects analysis
data_long <- data %>%
  arrange(participant_id, round) %>%
  group_by(participant_id) %>%
  mutate(order = row_number())

# Visualize learning effects
ggplot(data_long, aes(x = order, y = node_reached, color = treatment)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(alpha = 0.5) +
  labs(title = "Learning Effects Across Rounds",
       x = "Order of Rounds",
       y = "Node Reached") +
  theme_minimal()
```

