---
title: "Demo"
author: "Adam Hardaker"
date: "2024-06-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(lme4)
library(lmertest)
library(emmeans)
library(psych)
library(car)
library(gridExtra)
library(ggplot2)
```

loading the data
```{r}
file_path <- "data/ACTUAL_EXP_CREAMERY_QMEESS24.csv"

# Read the CSV file
df <- read_csv(file_path)
```

Cleaning data and using long format
```{r}
# Extract survey responses
survey_responses <- df %>%
  select(
    participant_id = `participant.id_in_session`,
    participant_label = `participant.label`,
    altruism = `centipede_game.4.player.altruism`,
    trust = `centipede_game.4.player.trust`,
    risk_tolerance = `centipede_game.4.player.risk_tolerance`,
    greed = `centipede_game.4.player.greed`,
    desire_to_win = `centipede_game.4.player.desire_to_win`
  ) %>%
  mutate(across(altruism:desire_to_win, ~as.numeric(as.character(.))))  # Convert to numeric

# Restructure the game data
restructured_data <- df %>%
  pivot_longer(
    cols = starts_with("centipede_game."),
    names_to = c("round", ".value"),
    names_pattern = "centipede_game\\.(\\d+)\\.(.+)",
    values_drop_na = TRUE
  ) %>%
  select(
    participant_id = `participant.id_in_session`,
    participant_label = `participant.label`,
    round,
    group_id = `group.id_in_subsession`,
    role = `player.id_in_group`, # indicates first-mover (1) or second mover (2)
    treatment = `group.treatment`,
    node_reached = `group.node`,
    payoff = `player.payoff`,
    treatment_order = `group.treatment_order`,
    
  ) %>%
  mutate(
    round = as.integer(round),
    group_id = as.integer(group_id),
    role = as.integer(role),
    node_reached = as.integer(node_reached),
    payoff = as.numeric(payoff),
  )

# Join game data with survey responses
final_data <- restructured_data %>%
  left_join(survey_responses, by = c("participant_id", "participant_label"))

# Sort the data by group_id, then round, then role
sorted_data <- final_data %>%
  arrange(group_id, round, role)

rm(restructured_data,final_data,df)

data<-sorted_data
# you can also save this if you wanna
```

## Analysis

Descriptive stats:
1. Sample size and demographics
2. Number of groups
3. Treatment-Specific Statistics
4. Round specific statistics
5. Treatment Order Effects
6. Personality Traits
7. Group dynamics
- frequency of games ended due to player 1 vs player 2 decisions
- average difference in payoffs between player 1 and player 2

```{r}
# Overall game statistics
overall_stats <- data %>%
  summarise(
    mean_node = mean(node_reached),
    median_node = median(node_reached),
    sd_node = sd(node_reached),
    mean_payoff = mean(payoff)
  )

# Treatment-specific statistics
treatment_stats <- data %>%
  group_by(treatment) %>%
  summarise(
    mean_node = mean(node_reached),
    median_node = median(node_reached),
    sd_node = sd(node_reached),
    mean_payoff = mean(payoff)
  )

# Round-specific statistics
round_stats <- data %>%
  group_by(round) %>%
  summarise(
    mean_node = mean(node_reached)
  )

# Personality traits summary
personality_stats <- data %>%
  select(altruism, trust, risk_tolerance, greed, desire_to_win) %>%
  describe()

# Correlation of traits with game outcomes
trait_correlations <- data %>%
  select(node_reached, payoff, altruism, trust, risk_tolerance, greed, desire_to_win) %>%
  cor()

# Print results
print(overall_stats)
print(treatment_stats)
print(round_stats)
print(personality_stats)
print(trait_correlations)

# You might also want to create visualizations, e.g.:
ggplot(data, aes(x = treatment, y = node_reached)) +
  geom_boxplot(varwidth=T, fill="plum", color="black") + 
    labs(title="Node Reached Grouped by Treatment",
         x="Treatment",
         y="Node Reached",
         fill="Treatment") +
  theme_minimal()

ggplot(data, aes(x = round, y = node_reached, color = treatment)) +
  geom_line(stat = "summary", fun = "mean", linewidth = 2) +
  labs(title = "Average Nodes Reached by Round and Treatment",
       y="Node Reached",
       x="Round",
       color="Treatment") +
  theme_minimal()
```

test normality
```{r}
shapiro_test <- shapiro.test(data$node_reached)
shapiro_test
```
Data significantly deviates from normal distribution

Some more normality stuff
```{r}
# Q-Q plot
qq_plot <- ggplot(data, aes(sample = node_reached)) +
  stat_qq() +
  stat_qq_line() +
  theme_minimal() +
  labs(title = "Q-Q Plot of Node Reached")

plot(qq_plot)

# Histogram
hist_plot <- ggplot(data, aes(x = node_reached)) +
  geom_histogram(binwidth = .5, fill = "orange", color="black", alpha = 1) +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  scale_x_continuous(breaks = seq(1, 6, by = 1)) +
  theme_minimal() +
  labs(title = "Histogram of Node Reached", x = "Node Reached", y = "Frequency")

plot(hist_plot)

library(ggplot2)
# Plot
g <- ggplot(data, aes(node_reached))
g + geom_density(aes(fill=factor(treatment)), alpha=0.4) + 
    labs(title="Node Reached Grouped by Treatment", 
         x="Node Reached",
         fill="Treatment")

# Plot
g2 <- ggplot(data, aes(treatment, node_reached))
g2 + geom_boxplot(varwidth=T, fill="plum") + 
    labs(title="Node Reached Grouped by Treatment",
         x="Treatment",
         y="Node Reached",
         fill="Treatment")
```
Doesn't look normal

## statistical tests

Use wide data for Friedman test
```{r}
data_wide <- data %>%
  group_by(participant_id, treatment) %>%
  summarize(avg_node_reached = mean(node_reached, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = treatment, values_from = avg_node_reached) %>%
  drop_na()

print(head(data_wide))

friedman_result <- friedman.test(as.matrix(data_wide[, c("control", "higher_fixed", "higher_random")]))
print(friedman_result)
```

Check learning effects
```{r}
# Prepare data for learning effects analysis
data_long <- data %>%
  arrange(participant_id, round) %>%
  group_by(participant_id) %>%
  mutate(order = row_number())

# Fit linear mixed model to test for learning effects
lmm_learning <- lmer(node_reached ~ order + treatment + (1|participant_id), data = data_long)
print("\nLinear Mixed Model for Learning Effects:")
print(summary(lmm_learning))

# Visualize learning effects
ggplot(data_long, aes(x = order, y = node_reached, color = treatment)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(alpha = 0.5) +
  labs(title = "Learning Effects Across Rounds",
       x = "Order of Rounds",
       y = "Node Reached") +
  theme_minimal()
```

